19Q1

### 2019-04-01 19Q1.Rmd
update Poz Venns, update slides for meeting, update unmatched histologies

# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_unmatched_na_histologies)
```

@@

## Matched data

```{r, eval=TRUE}
auc_matched <- read.delim("./data_munging/pharmacogenomics_raw_files/celline_stack.csv.gz", sep = ",")
auc_matched <- merge(ccl_histologies, auc_matched , by = "Cell_line", all.y = TRUE)
num_auc_per_db_matched <- auc_matched %>% count(Database)
ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line))
num_ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Database)
# only CCLs represetented in 2 or 3 databases were included
num_db_per_ccl_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Cell_line)
```

@@

auc_all_ccls <- bind_rows(auc_unmatched_wide, ccls_not_drugscreened)
auc_all_ccls$HasMutationData <- with(auc_all_ccls, ifelse(Broad_ID %in% maf_df$Broad_ID, 1, 0))
auc_all_ccls$HasCopyNumberData <- with(auc_all_ccls, ifelse(Broad_ID %in% cn_melt$Broad_ID, 1, 0))
auc_all_ccls$HasGeneExpressionData <- with(auc_all_ccls, ifelse(Broad_ID %in% ge_melt$Broad_ID, 1, 0))
auc_all_ccls$WasCRISPRScreened <- with(auc_all_ccls, ifelse(Broad_ID %in% crispr_meta$Broad_ID, 1, 0))
auc_all_ccls$InCCLE <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCCLE))
auc_all_ccls$InCTRP <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCTRP))
auc_all_ccls$InGDSC <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InGDSC))
auc_all_ccls[, -c(1:4)] <- apply(auc_all_ccls[, -c(1:4)], MARGIN = 2, FUN = as.numeric)
auc_all_ccls <- subset(auc_all_ccls, !(WasDrugScreened == 0 & WasCRISPRScreened == 0 & HasMutationData == 0 & HasGeneExpressionData == 0 & HasCopyNumberData == 0))
# write.csv.gz(auc_all_ccls, file = "./data_munging/pozdeyev2016/merged_auc_v20190327.3.csv.gz", row.names = FALSE, na = "NA")

@@

### Compar to TCGA

```{r}
cn_seg <- read.delim("./data_munging/19Q1/copy_number_20190328.csv.gz", sep = ",")
cn_seg <- merge(ccl_histologies, cn_seg, by = "Broad_ID", all.y = TRUE)
cn_seg_lung <- filter(cn_seg, Histology == "lung")
```

@@

# Denise's GCN dataframes (Deprected)

>>>>

ccl_histologies <- subset(ccl_histologies, !(Broad_ID %in% auc_unmatched_na_histologies$Broad_ID))
# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_unmatched_na_histologies)
```

@@

@@

auc_all_ccls <- bind_rows(auc_unmatched_wide, ccls_not_drugscreened)
auc_all_ccls$HasMutationData <- with(auc_all_ccls, ifelse(Broad_ID %in% maf_df$Broad_ID, 1, 0))
auc_all_ccls$HasCopyNumberData <- with(auc_all_ccls, ifelse(Broad_ID %in% cn_melt$Broad_ID, 1, 0))
auc_all_ccls$HasGeneExpressionData <- with(auc_all_ccls, ifelse(Broad_ID %in% ge_melt$Broad_ID, 1, 0))
auc_all_ccls$WasCRISPRScreened <- with(auc_all_ccls, ifelse(Broad_ID %in% crispr_meta$Broad_ID, 1, 0))
auc_all_ccls$InCCLE <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCCLE))
auc_all_ccls$InCTRP <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InCTRP))
auc_all_ccls$InGDSC <- with(auc_all_ccls, ifelse(WasDrugScreened == 0, 0, InGDSC))
auc_all_ccls[, -c(1:4)] <- apply(auc_all_ccls[, -c(1:4)], MARGIN = 2, FUN = as.numeric)
auc_all_ccls <- subset(auc_all_ccls, !(WasDrugScreened == 0 & WasCRISPRScreened == 0 & HasMutationData == 0 & HasGeneExpressionData == 0 & HasCopyNumberData == 0))
# write.csv.gz(auc_all_ccls, file = "./data_munging/pozdeyev2016/merged_auc_v20190401.4.csv.gz", row.names = FALSE, na = "NA")

@@

```{r}
auc_all_ccls <- read.delim("./data_munging/pozdeyev2016/merged_auc_v20190401.4.csv.gz", sep = ",")

sum(auc_all_ccls$HasCopyNumberData)
sum(auc_all_ccls$HasMutationData)
sum(auc_all_ccls$HasGeneExpressionData)
sum(auc_all_ccls$WasDrugScreened)

pgx_ccls <- filter(auc_all_ccls, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all_ccls, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all_ccls, WasCRISPRScreened == 0 & WasDrugScreened == 0)

ge_ccls <- filter(auc_all_ccls, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all_ccls, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all_ccls, HasMutationData == 1)

all_omics_ccls <- intersect(intersect(unique(ge_ccls$Cell_line), unique(cn_ccls$Cell_line)), unique(mut_ccls$Cell_line))
sum(filter(auc_all_ccls, Cell_line %in% all_omics_ccls)$WasDrugScreened)

ccle_ccls <- filter(auc_all_ccls, InCCLE == 1)
ctrp_ccls <- filter(auc_all_ccls, InCTRP == 1)
gdsc_ccls <- filter(auc_all_ccls, InGDSC == 1)

auc_all_ccls %>% count(Histology)
```

```{r}
test <- mut_ccls[, 1:20]
test2 <- test %>% count(Broad_ID)
```

### Summary figures

```{r, eval=FALSE}
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_pgx_crispr_overlap <- venn.diagram(list("CRISPR" = unique(crispr_ccls$Cell_line), "Drug Screen" = unique(pgx_ccls$Cell_line), "Unscreened" = unique(unscreened_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened via small-molecule and CRISPR", main.fontface = "bold", lty = rep("blank", 3), fill = c("mediumpurple1", "turquoise4", "slategray3"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/venn_pgx_crispr_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_pgx_crispr_overlap)
dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_ccl_overlap <- venn.diagram(list("CCLE" = unique(ccle_ccls$Cell_line), "CTRP" = unique(ctrp_ccls$Cell_line), "GDSC" = unique(gdsc_ccls$Cell_line)), resolution = 5000, main = "Overlap of cancer cell lines (CCLs) screened in each dataset", main.fontface = "bold", lty = rep("blank", 3), fill = c("springgreen4", "steelblue4", "turquoise4"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
# png("./gcn_denise/pozdeyev_unmatched/pozdeyev_venn_all_ccl_unmatched.png", res = 500, units = "in", width = 6, height = 6)
# grid.newpage()
# grid.draw(venn_ccl_overlap)
# dev.off()

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")
venn_omics_overlap <- venn.diagram(list("Gene Expression" = unique(ge_ccls$Cell_line), "Copy Number" = unique(cn_ccls$Cell_line), "Mutation" = unique(mut_ccls$Cell_line)), resolution = 5000, main = "Overlapping omic annotations in cancer cell lines (CCLs)", main.fontface = "bold", lty = rep("blank", 3), fill = c("yellow2", "orange", "tomato"), alpha = c(0.5, 0.5, 0.5), cat.pos = c(-30, 30, 180), cat.dist = c(0.01, 0.001, 0.001), cat.cex = 1.5, cex = 1, filename = NULL)
png("./gcn_denise/pozdeyev_unmatched/pozdeyev_venn_omics_overlap.png", res = 500, units = "in", width = 6, height = 6)
grid.newpage()
grid.draw(venn_omics_overlap)
dev.off()

num_ccl_per_hist_unmatched_bar <- ggplot(data = unique(select(auc_all_ccls, Cell_line, Histology))) +
  geom_bar(aes(x = Histology, fill = Histology), alpha = 0.7) +
  guides(fill = "none") +
  scale_y_continuous(breaks = seq(0, 300, by = 25), labels = seq(0, 300, by = 25)) +
  labs(y = "Number of cell lines", title = "Number of unique cell lines per histology across all cell lines") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
num_ccl_per_hist_unmatched_bar
# ggsave("./gcn_denise/pozdeyev_unmatched/pozdeyev_ccl_per_histology_unmatched.png", num_ccl_per_hist_unmatched_bar, device = "png", dpi = 450, width = 12, height = 6, units = "in")
```

### Compare to TCGA
Quality check: cluster TCGA samples and cell lines—verify that same tumor types cluster together rather than clustering by cell lines and tumor samples

Calculate Pearson correlations between data sets and create a heatmap (also look at Barretina et al. supplement for PCA and mutation load correlations)

```{r, eval=FALSE}
cn_seg <- read.delim("./data_munging/19Q1/copy_number_20190328.csv.gz", sep = ",")
tcga_pancan <- read.delim("~/Desktop/GDC-PANCAN.masked_cnv.tsv.gz", sep = "\t")

ggplot(cn_seg, aes(x = log2(Segment_Mean/2))) + 
  geom_density()

ggplot(tcga_pancan, aes(x = value)) + 
  geom_density()
```

@@

# (OLD) Pozdeyev et al (2016) matched data

```{r, eval=FALSE}
auc_matched <- read.delim("./data_munging/pharmacogenomics_raw_files/celline_stack.csv.gz", sep = ",")
auc_matched <- merge(ccl_histologies, auc_matched , by = "Cell_line", all.y = TRUE)
num_auc_per_db_matched <- auc_matched %>% count(Database)
ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line))
num_ccl_per_db_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Database)
# only CCLs represetented in 2 or 3 databases were included
num_db_per_ccl_matched <- unique(select(auc_matched , Database, Cell_line)) %>% count(Cell_line)
```

# (OLD) Denise's GCN dataframes
